---
title: "PCA"
author: "Ana"
date: "14/6/2020"
output:
  word_document: default
  html_document: default
---
1.PCA with FactoMineR function
```{r}
library(dplyr)
library (ggplot2)
library(FactoMineR)

final_data <- read.csv("C:/Users/Ana/Desktop/final_data.csv")
sapply(final_data, class)
#remove factor and responsive vble
fd_for_pca <- select(final_data, 
                     -countries_en, 
                     -nutriscore_grade,
                     -pnns_groups_2,
                     -nutriscore_score)

head(final_data, n=9)
#data for PCA 36556 - 8 vbles


#PCA con funcion
pca_final_data <- prcomp(fd_for_pca, center = TRUE, scale. = TRUE)
summary(pca_final_data)
biplot(pca_final_data, choices=1:2)
library(factoextra)
fviz_pca_var(pca_final_data,
             col.var = "contrib", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     
             )
#compute standard deviation for each component
standard_deviation <- pca_final_data$sdev
#compute variance
var <- standard_deviation^2
#variance of 4 first PCA
var[1:6]
#proportion of variance explained
prop_varex <- var/sum(var)
prop_varex[1:8]
plot(prop_varex, xlab = "Principal Component",
             ylab = "Proportion of Variance Explained",
             type = "b")


library(corrplot)
var <- get_pca_var(pca_final_data)
corrplot(var$cos2, is.corr=FALSE)
fviz_cos2(pca_final_data, choice = "var", axes = 1:2)

```
according to last elbow rule, the first 4 PCA are significant, retaining a total of 73% of inertia

The most well represented feature is carbohydrates along sugar as these two are usually representative as the same nutrient in foods. 


the least represented dimension is salt, as the value registered in the dataset corresponds to a 100g ration of the product, and the retio of this ingredient in that amount of food is usually not representative.


2.PCA function with categorial dimensions as sumplementary
```{r}
#PCA con factors as supplementary to plot in same factorial plane
fd1 <- select(final_data,-nutriscore_score)
pca1 <- PCA(fd1,quali.sup=c(1,3,5)) 

fviz_pca_biplot(pca1, label = "var", habillage=fd1$nutriscore_grade)
fviz_pca_biplot(pca1, label = "var", habillage=fd1$countries_en)

#sampling for a better visualization of countries as suplementary 
sampleindex <- sample(1:nrow(fd1), 500)
fdsample <- fd1[sampleindex, ]
pca2 <- PCA(fdsample,quali.sup=c(1,3,5))
fviz_pca_biplot(pca2, label = "var", habillage=fdsample$nutriscore_grade)
fviz_pca_biplot(pca2, label = "var", habillage=fdsample$countries_en)

```
high nutrition score grades are mostly foods with high protein values

```
```{r}
```

