---
title: "linear_regression"
author: "Ana"
date: "17/6/2020"
output: html_document
---
```{r}

library("matrixStats")
library("ggplot2")
library("gridExtra")
library("ggpubr")
library("DMwR")
library("plsdepot")
library("FactoMineR")
library(chemometrics)

```
```{r}
#data 
mydata <- read.csv("C:/Users/Ana/Desktop/final_data.csv")
sapply(mydata, class)
head(mydata, n=8)

#split data 
set.seed(1876)
test.index = sample(1:nrow(mydata), nrow(mydata)*0.1)
#data.nograde = subset(mydata, select= -nutriscore_grade)
data.test = mydata[test.index,]
data.train = mydata[-test.index,]

class(mydata)

library(dplyr)
library(FactoMineR)

target.vble <- select(mydata, nutriscore_score)
explanat.vbles <- select (mydata, -nutriscore_score)

x.target <- data.matrix(target.vble)
x.explanat <- data.matrix(explanat.vble)
head(x.explanat)
```


1. multiple linear regression 
```{r}
#remove NA if any
sapply(mydata, function(x) sum(is.na(x)))

#model1: keeping only macronutrients
model1 <- glm(formula = nutriscore_score ~ proteins_100g + carbohydrates_100g + fat_100g, data = data.train)
summary(model1)
mse1 <- mean(model1$residuals^2)
mse1

#model2: keeping all explanatory features
model2 <- glm(nutriscore_score ~ proteins_100g + carbohydrates_100g + fat_100g + countries_en + additives_n + nova_group + nutriscore_grade + pnns_groups_2 + sugars_100g + fiber_100g + salt_100g, data = data.train, family = gaussian)
summary(model2)
mse2 <- mean(model2$residuals^2)
mse2

#model3: keeping only nutritional features, remove countries_en, pnns_group
model3 <- glm(nutriscore_score ~ proteins_100g + carbohydrates_100g + fat_100g + additives_n + nova_group + nutriscore_grade + sugars_100g + fiber_100g + salt_100g, data = data.train, family = gaussian)
summary(model3)
mse3 <- mean(model3$residuals^2)
mse3
confint(model3, conf.level=0.9)

plot(data.train$salt[data.train$nutriscore_grade=="a"], data.train$nutriscore_score[data.train$nutriscore_grade=="a"], col=2)
points(data.train$salt[data.train$nutriscore_grade=="b"], data.train$nutriscore_score[data.train$nutriscore_grade=="b"], col=3)
points(data.train$salt[data.train$nutriscore_grade=="c"], data.train$nutriscore_score[data.train$nutriscore_grade=="c"], col=4)
points(data.train$salt[data.train$nutriscore_grade=="d"], data.train$nutriscore_score[data.train$nutriscore_grade=="d"], col=5)
points(data.train$salt[data.train$nutriscore_grade=="e"], data.train$nutriscore_score[data.train$nutriscore_grade=="e"], col=6)

plot(data.train$nova_group[data.train$nutriscore_grade=="a"], data.train$nutriscore_score[data.train$nutriscore_grade=="a"], col=2)
points(data.train$nova_group[data.train$nutriscore_grade=="b"], data.train$nutriscore_score[data.train$nutriscore_grade=="b"], col=3)
points(data.train$nova_group[data.train$nutriscore_grade=="c"], data.train$nutriscore_score[data.train$nutriscore_grade=="c"], col=4)
points(data.train$nova_group[data.train$nutriscore_grade=="d"], data.train$nutriscore_score[data.train$nutriscore_grade=="d"], col=5)
points(data.train$nova_group[data.train$nutriscore_grade=="e"], data.train$nutriscore_score[data.train$nutriscore_grade=="e"], col=6)

#create dummy variable/indicator
levels(data.train$nutriscore_grade)
mean(data.train$nutriscore_grade=="a")
mean(data.train$nutriscore_grade=="b")
mean(data.train$nutriscore_grade=="c")
mean(data.train$nutriscore_grade=="d")
mean(data.train$nutriscore_grade=="e")
```
```
```{r}
mse <- mean(model3$residuals^2)
install.packages("Metrics")
library(Metrics)
pd <- predict(model3 , data.train)
summary(pd)

mse
#[1] 168.7736
```

model4
```{r}
#model 4 with log target variable
model4 <- glm(log(nutriscore_score) ~ proteins_100g + carbohydrates_100g + fat_100g + additives_n + nova_group  + sugars_100g + fiber_100g + salt_100g, data = data.train, family = gaussian)
summary(model4)
model4$coefficients
confint(model4, conf.level=0.9)
mse4 <- mean(model4$residuals^2)
mse4
#[1] 0.5182961
#we keep it 
```

model5
```{r}
model5 <- glm(log(nutriscore_score) ~ proteins_100g + carbohydrates_100g + fat_100g + additives_n + poly(nova_group,2)  + sugars_100g + fiber_100g + salt_100g, data = data.train, family = gaussian)
summary(model5)
model5$coefficients
confint(model5, conf.level=0.9)
mse5 <- mean(model5$residuals^2)
mse5
#[1] 0.5176696
```

model6
```{r}
model6 <- glm(log(nutriscore_score) ~ proteins_100g + carbohydrates_100g + fat_100g + additives_n + nova_group  + sugars_100g + poly(fiber_100g,2) + salt_100g, data = data.train, family = gaussian)
summary(model6)
model66 <- lm(log(nutriscore_score) ~ proteins_100g + carbohydrates_100g + fat_100g + additives_n + nova_group  + sugars_100g + poly(fiber_100g,2) + salt_100g, data = data.train, family = gaussian)
summary(model66)
model6$coefficients
confint(model6, conf.level=0.9)
mse6 <- mean(model6$residuals^2)
mse6
#[1] 0.5168053

head (data.train, n=9)
```
we keep model 3

2. ridge regression
```{r}
library(stats)
library(broom)
library(glmnet)

x = data.matrix(data.train)
y_train = data.train$nutriscore_score
alpha = 0
lambdas <- 10^seq(3, -2, by = -.1)

target.vble <- select(mydata, nutriscore_score)
explanat.vbles <- select (mydata, -nutriscore_score, -countries_en, -nutriscore_grade, pnns_groups_2)

x.target <- data.matrix(target.vble)
x.explanat <- data.matrix(explanat.vble)
head(x.explanat)

ridge_reg = glmnet(x.explanat, x.target, nlambda = 25, alpha = 0, lambda = lambdas)
summary(ridge_reg)

lasso_reg <- cv.glmnet(x.explanat, x.target, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)


class(explan2)
class(x.explanat)

